---
title: "ZED Project"
author: "Piotr Falkiewicz 122563"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    keep_md: yes
    toc: true
    toc_float: true
    
---

```{r setup_libs, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache.lazy = FALSE)
library(dplyr)
library(ggplot2)
library(DT)
library(data.table)
library(knitr)
library(plotly)
library(corrplot)
library(reshape2)
library(caret)
library(randomForest)


prettyTable1 <- function(table_df, round_columns=numeric(), round_digits=2) {
    DT::datatable(table_df, style="bootstrap", rownames = TRUE, editable = FALSE,
    width = 500,
    options = list(dom = 'tp')) %>%
    formatRound(round_columns, round_digits)
}

prettyTable2 <- function(table_df, round_columns=numeric(), round_digits=2) {
    DT::datatable(table_df, style="bootstrap", rownames = TRUE, editable = FALSE,
    width = 400,
    options = list(dom = 't')) %>%
    formatRound(round_columns, round_digits)
}
```

```{r load_data, include=FALSE, cache = TRUE}
forbidden_columns <- c( "UNK", "UNX", "UNL", "DUM", "N", "BLOB", "ALA", "ARG", "ASN", "ASP", "CYS", "GLN", "GLU", "GLY", "HIS", "ILE", "LEU", "LYS", "MET", "MSE", "PHE", "PRO", "SEC", "SER", "THR", "TRP", "TYR", "VAL", "DA", "DG", "DT", "DC", "DU", "A", "G", "T", "C", "U", "HOH", "H20", "WAT")

rawdata <- fread("all_summary.csv")
set.seed(19)
```

#Introduction 
```{r zero_part,message=FALSE, echo=FALSE, warning = FALSE, cache = TRUE}
data_zero <- rawdata %>%
  filter(!is.na(res_name)) %>%
  filter(!res_name %in% forbidden_columns) %>%
  select_if(is.numeric)

num_stats <- t(sapply(data_zero, function(x){list(round(min(x, na.rm=TRUE),2),round(max(x, na.rm=TRUE),2),round(mean(x, na.rm=TRUE),2),round(median(x,na.rm=TRUE),2))}))
colnames(num_stats) <- c("Min", "Max", "Mean", "Median")
```


Report stands for analyse of Protein Bank Data. 



The input dataset consists of `r dim(rawdata)[1]` observations and `r dim(rawdata)[2]` attributes.

At the start of analyse, dataset was cleared from observations that didn't belong to 50 most popular classes of protein.
Skipping non-compelete rows was the way of handling NA values.

Statistics of numeric attributes can be spotted in the tabel below.

`r prettyTable1(num_stats)`

# 'res_name' class analysis 

```{r first_part_analysis, echo=FALSE}
data_first <- rawdata %>%
  select(c(res_name, local_res_atom_non_h_count, local_res_atom_non_h_electron_sum, local_res_atom_non_h_count, dict_atom_non_h_count, local_res_atom_non_h_electron_sum, dict_atom_non_h_electron_sum)) %>% 
  filter(!is.na(res_name)) %>%
  filter(!res_name %in% forbidden_columns) 

dataset_first <- data_first[complete.cases(data_first),]

top_50_classes <- dataset_first %>% 
  group_by(res_name) %>%
  summarise(count = n()) %>% 
  arrange(desc(count)) %>%
  head(50)

dataset_first_processed <- filter(dataset_first, res_name %in% top_50_classes$res_name)
ds_first_processed_no_nan <- dataset_first_processed[complete.cases(dataset_first_processed),]

plot1 <- ggplot(ds_first_processed_no_nan, aes(x = local_res_atom_non_h_count)) + geom_density() + theme_bw()

plot2 <- ggplot(ds_first_processed_no_nan, aes(x = local_res_atom_non_h_electron_sum)) + geom_density() + theme_bw()

ds_first_atoms_diff <- ds_first_processed_no_nan %>% 
  filter(local_res_atom_non_h_count != dict_atom_non_h_count) %>%
  mutate(difference = abs(local_res_atom_non_h_count - dict_atom_non_h_count)) %>%
  group_by(res_name) %>%
  summarise(total_diff = sum(difference), count = n()) %>% 
  arrange(desc(total_diff)) %>%
  head(10)

ds_first_electrons_diff <- ds_first_processed_no_nan %>% 
  filter(local_res_atom_non_h_electron_sum != dict_atom_non_h_electron_sum) %>%
  mutate(difference = abs(dict_atom_non_h_electron_sum-local_res_atom_non_h_electron_sum)) %>%
  group_by(res_name) %>%
  summarise(total_diff = sum(difference), count = n()) %>% 
  arrange(desc(total_diff)) %>%
  head(10)

```

The top 50 most popular values of class 'res_name' are shown in table below:
`r prettyTable1(top_50_classes)`

Distribution of number of atoms and electrons has been shown on charts:

a) For atoms
`r ggplotly(plot1)`

b) For electrons:
`r ggplotly(plot2)`

That part of analysis contains 10 classes with biggest difference between 

a) For atoms
`r prettyTable2(ds_first_atoms_diff)`

b) For electrons
`r prettyTable2(ds_first_electrons_diff)`

# Correlation between variables
```{r second_part_correlation, message=FALSE, echo=FALSE, warning = FALSE, cache = TRUE}
ds_second <- rawdata %>% select_if(is.numeric)
ds_second2 <- ds_second 
ds_second_cor <- ds_second2[complete.cases(ds_second2),]
ds_second_cor_final <- round(cor(ds_second_cor),2)
ds_second_cor_melted <- melt(ds_second_cor_final) %>% arrange(value)

corr_plot <- ggplot(ds_second_cor_melted, aes(x=Var1, y=Var2, fill=value)) + 
   geom_tile() + 
   geom_tile(color = "white") +
  
   scale_fill_gradient2(low = "red", high = "green", mid = "yellow", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  
   theme(
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
   ) 

#ggplotly(corr_plot)

```
This section shows heatmap of correlation between variables
`r ggplotly(corr_plot)`

Highest value of correlation can be noticed for columns representing features of shape or dimensions of protein.

# "part_01*"" columns analysis

That part consist of distribution figures for columns, with names starting with 'part01'. The distribution is represented using density plots.

```{r third_part_columns, echo=FALSE, warning=FALSE, cache=TRUE}

dataset_third <- rawdata %>% 
select(starts_with("part_01"))

for(colName in names(dataset_third)) {
  currentCol <- dataset_third[[colName]]
  currentMean <- mean(currentCol[!is.na(currentCol)])
  currentPlot <- ggplot(dataset_third, aes(x = currentCol)) + 
  xlab(colName) +
  ylab("Density") +
  geom_density(kernel = "gaussian", fill = "olivedrab3") + 
  theme_minimal() +
  geom_rug(sides="b", color="black", size = 1, aes(x=currentMean, y=0)) + 
  geom_text(aes(label=round(currentMean, 2), x=currentMean, y=0), hjust=0.5, vjust=-1, color="black")
  print(currentPlot)
}


```

# Predicting number of electrons/atoms - linear regression

```{r forth_part_electrons_atoms_regression, echo=FALSE, warning=FALSE, cache=TRUE}
ds_forth <- rawdata %>% select(res_name, local_res_atom_non_h_count, local_res_atom_non_h_electron_sum,resolution,FoFc_mean,FoFc_std,FoFc_square_std,FoFc_min,FoFc_max) %>% filter(res_name %in% top_50_classes$res_name)
ds_forth <- ds_forth[complete.cases(ds_forth),]

# Elektrony
ds_forth_electrons <- ds_forth %>% select(-c(local_res_atom_non_h_count,res_name))
lm_e <- train(local_res_atom_non_h_electron_sum ~., data = ds_forth_electrons, method = "lm")

# Atomy
ds_forth_atoms <- ds_forth %>% select(-c(local_res_atom_non_h_electron_sum,res_name))

lm_a <- train(local_res_atom_non_h_count ~., data = ds_forth_atoms, method = "lm")

```

The experiment included attemp to create linear regression model for number of atoms(local_res_atom_non_h_count) and electrons(local_res_atom_non_h_electron_sum).

The quality of linear regression for atoms is `r round(lm_a$results$RMSE,2)` for RSME and `r round(lm_a$results$Rsquared,2)` for R^2.
The quality of linear regression for electrons is `r round(lm_e$results$RMSE,2)` for RSME and `r round(lm_e$results$Rsquared,2)` for R^2.

# Predicting res_name class value - classifier
```{r fifth_part_res_name_classifier, echo=FALSE, warning=FALSE, cache=TRUE}
ds_fifth <- rawdata %>% select(res_name, resolution,FoFc_mean,FoFc_std,FoFc_square_std,FoFc_min,FoFc_max) %>% filter(res_name %in% top_50_classes$res_name)
ds_fifth <- ds_fifth[complete.cases(ds_fifth),]
ds_fifth$res_name <- factor(ds_fifth$res_name, labels=c(1:length(table(ds_fifth$res_name))))

splitRatio <- 0.7
ds_fifth_train_id <- createDataPartition(ds_fifth$res_name, p = splitRatio, list = FALSE)
ds_fifth_train <- ds_fifth[ds_fifth_train_id,]
ds_fifth_test <- ds_fifth[-ds_fifth_train_id,]

ctrlC <- 5
ctrlV <- 10

ctrl <- trainControl(method = "repeatedcv", number = ctrlC, repeats = ctrlV)

nTree <- 13

ds_fifth_fit <- train(res_name ~ ., data = ds_fifth_train, method = "rf", trControl = ctrl, ntree = nTree)
df_fifth_predictions <- predict(ds_fifth_fit, newdata = ds_fifth_test)

cm <- confusionMatrix(data = df_fifth_predictions, ds_fifth_test$res_name)
```

Classifier of res_name class was based on Random Forest algorithm. 

First data was cleared from NaN cases and then splitted into train and test batches with ratio `r splitRatio`.
The train control of model consists of `r ctrlC` parts and `r ctrlV` repeats.

The quality measures for test batch are `r round(cm$overall[1],2)` for accuracy and `r round(cm$overall[2],2)` for Kappa score.

